\bf Mayo Pharmacy Optimization Case Study\\ \large Using Discrete Event Simulation

### 2019-12-13 Mayo/Final Draft/Draft.tex
Katies Edits

The use of prescription medication is becoming increasingly commonplace in the United States. In 2013, the average number of prescriptions taken per person was $12.2$, up from $9$ in 1998, a $136\%$ increase over 15 years \cite{georgetown}\cite{statistica}. With this trend in prescription usage continuing to rise, pharmacies are becoming an ever greater presence in the lives of most Americans. As such, pharmacies need to adapt to deal with increasing numbers of orders, as well as increasing the speed and accuracy at which they fill orders in order to remain competitive in an increasingly saturated landscape of pharmacies.  Consequently, pharmacies must adapt their methodologies and staffing requirements in order to optimize efficiency and cost. Obviously, testing different optimization schemas is both expensive and time consuming, so pharmacies have begun to consider computer simulation approaches to optimization as a more cost effective solution.\\\hfill\\
In this case study, we are tasked with both modeling and optimizing the workflow for a prototype pharmacy based on samples of data regarding task duration and incoming order frequency. In addition to the sample data for the pharmacy, we are also given a brief overview of the basic order fulfillment process. Using this information, we shall construct a discrete event simulation model to analyze our so called ``control" case, i.e. the pharmacy described by the basic information. Once this simulation is running and tested, we shall attempt to improve the efficiency over the base case. In order to achieve this goal, we shall attack the optimization on two fronts.

@@

With these optimizations, we hope to develop a sense of what changes could increase the efficiency of the pharmacy, as well as an idea of what variables could be manipulated to further increase the accuracy of our simulation, and improve the pharmacy's function.

@@

In our prototypical example, there are two types of prescription orders that the pharmacy receives, oral and IV, which are handled by two different classes of employee, pharmacists and technicians, who are constrained as to which task they may perform due to legal regulations. In our prototype pharmacy, the proposed process is as follows: As the pharmacy receives orders, they are entered into the system by a technician. Once the order is in the system, it is verified by a pharmacist, and passed to the appropriate area to be prepared by technicians. Upon being filled, the order is again verified by a pharmacist, and then dispensed to the consumer by a technician. As such, we see that there are six basic tasks in the workflow tied to the two different types of orders: order entry, entry verification, oral or IV preparation, prepared order verification, and distribution through the window. As noted before, regulations exist to limit what tasks non-pharmacist level employees may perform. These regulations dictate that of the tasks, the verification tasks may only be performed by the pharmacists, while the remaining tasks can be performed by both pharmacists and technicians. A breakdown of this hierarchy follows in Table \ref{table:tasks}, with pharmacist specific tasks highlighted in red. It should also be noted that in our control case, there exists an additional sub-type of technician, whose sole purpose is to prepare IV orders in an isolated room. This will be explored in greater detail below.

@@

\begin{tabular}{|c||c|c|}

@@

Order Entry & Order Entry &\\\hline
\cellcolor{red!25}Entry Verification && \\\hline

@@

\cellcolor{red!25}Preparation Verification &&\\\hline
Distribution & Distribution&\\\hline

@@

With this task breakdown in hand, we consider a graphical representation of the path an order takes within the pharmacy, specifically linked to the tasks described above. Looking at Figure \ref{fig:flowchart}, we see that the process of filling an order is fairly linear, with the only fork in the workflow occurring after the entry verification step. This fork is due to the necessity of establishing an order's type (Oral vs. IV) so it may be put in the appropriate filling queue. In our control pharmacy, all IV orders are filled in an isolated room by an ``IV Technician" who only prepares IV prescriptions, and does not assist with any other tasks in the workflow. 

@@

\caption{Pharmacy workflow by task position noting which tasks are pharmacist only.}

@@

For our control case, we operate under the assumption that no tasks in this process have any issues with failure or delays. This is obviously a simplified case, but we shall construct the simulation in such a way that including it would be quite simple to do.  With this workflow defined, we consider the next step of building our simulation, identifying the durations of the different tasks.

@@

As noted above, there are six pivotal tasks that we must establish the duration of in order to effectively perform an analysis. In addition to these tasks, we are also aware of the fact that the times between incoming oral and IV orders come from different distributions. As part of the supplied data, we have samples of the completion times of each of these eight different categories, ranging in size from 34 to 60 observations. Using each of these samples, we shall construct Cullen and Frey graphs to attempt to isolate the correct distributions. These graphs will take the skewness and kurtosis of our task time samples and plot a point in the $skew^2/kurtosis$ plane which can then be mapped to potential distributions based on theoretical values for each distribution. With this set of hypothetical distributions in hand, we shall fit the data and test the goodness of fit statistic to identify the best fit.

@@

Looking at the histogram of order entry data in Figure \ref{fig:orderEntryHistCullen}, it appears that the data comes from a normal population, and the Cullen and Frey plot supports this. For completeness, we shall test this sample as being from a normal distribution as well as a gamma and log-normal distribution, as these distributions also appear to be good potential fits in the Cullen and Frey graph. Once the data has been fit with each of these distributions, $R$ is used to compute a goodness of fit statistic, the AIC, for each model. The AIC is a sort of score for a model, which weights the accuracy of a fitted model against its own complexity. By construction, this score decreases for better models, and increases when models become too complex. This prevents over-fitting and reliance on sample data values. With this in mind, we consider the AIC scores for each of the three potential distributions for order entry below in Table \ref{table:entryGOF}

@@

& Normal Fit & Gamma Fit & log-normal Fit \\\hline

@@

As highlighted in Table \ref{table:entryGOF}, by treating this sample as part of a larger normal population, we produce the smallest AIC score, implying the best fit. As such, we shall treat this as the correct distribution for sampling purposes in our simulation. The relevant distribution statistics follow,

@@

To construct the simulation of the pharmacy, the programming language C++ was used due to its speed, as well as object oriented structure. This structure allowed for maximal customization of our data, enabling the collection of many different statistics, which we report below. To build the simulation, two unique class objects were written to house the details required for a given prescription, as well as each of the employees. Within these classes relevant structures for computing the idle times and other important statistics are included and updated throughout the simulation. An outline of the pseudo-code for the simulation follows:

@@

It is important to note here the job assignment of the base case. In this model, when job assignment begins for each employee, each non-empty task queue is put into a pool and randomly chosen with no priority. So, a pharmacist has an equal probability of doing a pharmacist only task or a technician level task. Changing this assignment will become a cornerstone of our optimization efforts. To further remove any bias in the model, the idle worker pool is shuffled randomly before each task assignment step in the simulation. This will help to alleviate any artificial effects on worker idle times.

@@

We see that we expect just shy of $100$ oral prescriptions, and just shy of $50$ IV orders per day. This indeed aligns with the expected values given the exponential distribution fit that was assigned to them for time between orders.\\\hfill\\
Next, we consider the average queue lengths for each task in the workflow,

@@

Next, we consider arguably the most valuable statistic to a pharmacy from a business standpoint, the number of filled orders. 

@@

We see here that we expect to fill about $70\%-75\%$ of all incoming orders. To see whether there is room for improvement in this number, we consider whether the amount of time for orders from start to finish, the ``throughput", as well as the amount of time the order sits idle leave room for improvement.

@@

While knowing how long and how fast the orders are moving in the model is important, it is equally important for the pharmacy to measure the efficiency and use of its employees. Below, we see the estimations for how long each type of employee is idle during the simulations, expressed as a percent of the total time.

@@

We see that the results of this simulation point to the verification queues as being congestion points. Notably, these are the two tasks in the workflow that are pharmacist specific. This congestion then is likely due to the method by which workers received their job assignments. With the pharmacists choosing from all possible tasks equally, they can take work from the technicians, who cannot in return work from the verification tasks. This aligns with the estimated idle times from above, as we found the pharmacists to be idle significantly less than the technicians. Reducing this congestion will be the goal of our first optimization.

@@

With the base case of the model constructed and analyzed, we begin the first of our two pronged optimization efforts, optimizing the job choice logic to increase throughput speed, and reduce idle time in the model.
\subsection*{First Optimization: Lazy Pharmacists}
\addcontentsline{toc}{subsection}{Lazy Pharmacists}
Given that the verification tasks are causing congestion in the base model, we consider the effects of changing the job assignment method in the simulation to the following.

@@

Under this scheme, the pharmacists will only do pharmacist specific tasks, leaving the technicians to do the remainder of the tasks. The hope is that by removing the pharmacists from the technician tasks that not only will the technician idle times be reduced, but also the lengths of the verification queues will be reduced as well. Running the simulation with the same parameters as the base case, under the new job choice logic yields some unexpected results. Firstly, the average queue lengths are much different. Figure \ref{fig:basevprof} shows the differing results.

@@

\caption{Base case (left) vs. Lazy Pharmacist (Right) average queue lengths}

@@

Under the original schema, all queues were essentially empty except for the verification queues, while under our new schema all of the non-verification queues show increases, most specifically the first and last steps of the fill process. Indeed we find that the average queue length for our new case is slightly larger than before,
\[\overline{Length}_{Base}=2.648<2.898=\overline{Length}_{Lazy Pharmacist}\]

@@

\[\overline{Idle}_{Base}=14.475\% < 21.627\%=\overline{Idle}_{Lazy Pharmacist}\]
Looking more closely into this idle time data, we see that the average idle time for a pharmacist in the new scheme is approximately $67\%$, which is clearly not an effective use of their time. Before looking to see if reducing the number of pharmacists in the model would help to reduce this idle percentage, we first consider the filled order results under this scheme. Figure \ref{fig:baseprofFill} shows us a comparison in throughput.

@@

\caption{Base case (left) vs. Lazy Pharmacist (Right) average number of filled orders}

@@

We see that in fact this method has reduced both the number of oral and IV orders that are filled over the duration of the simulation. It is likely not worth exploring further in terms of staffing optimizations as an ``over-staffed" run yielded worse results. As such, we seek to find a better process optimization.

@@

Given that forcing the pharmacists to only pull from the verification tasks is hurting the efficiency of the model, we consider a new job assignment logic, choosing the most full task queue. So, a pharmacist will choose to take their next task from the largest of all five possible queues, while technicians may choose from the three technician level queues. This method should help to ensure that both employees and orders are moving as efficiently as possible as it favors maintaining the smallest possible net queue sizes. Running the simulation with the same parameters as the base case, under the new job choice logic yields results that are more aligned with what we would expect. Figure \ref{fig:basevlong} shows the comparison on average queue lengths between the base case and the longest queue method.

@@

We see that under this new system, all of the queues (except IV fill) have non-zero means, but that they are all relatively small compared to the mean of approximately 10 for the base entry verification length. Looking deeper into the data, we see that the average queue length is reduced from $2.648$ in the base case to $1.389$ in this new method. In addition to this, average idle time for the workers is reduced from $14.475\%$ in our base model to $13.395\%$. While these changes may seem somewhat small, consider now the mean throughput times between the two models.

@@

We see that by choosing the longest queue as the next task, the average throughput time for an order is reduced by a factor of almost $30\%$. As we would expect, this also helps to reduce the average wait time for orders, as well as the idle times of the employees.  While this method is clearly superior to the base case, we consider finally a combination of the Lazy Pharmacist and Longest Queue methods.
\subsection*{Third Optimization: Smart Method}
\addcontentsline{toc}{subsection}{Smart Method}
As our final job assignment optimization we consider a sort of mixture between choosing the longest queue and relegating the pharmacists to the verification queues. Namely, under this schema, the pharmacist will first check to see if any of the verification queues are non-empty, and if so, they will choose the largest one. Otherwise, they choose the largest from the remaining queues. The technicians remain unchanged from before, always choosing a task from the largest queue.

@@

Again we implement this model with the basic parameters from before, getting again better results than before. As before, we consider the comparison between the base case and the smart method queue lengths to begin to test whether we have achieved optimization.

@@

\caption{Base Case (left) vs. Smart Method (Right) average queue lengths}

@@

As with the longest queue method we see a significant reduction in the average queue length, going from $2.6475$ in the base case to $1.1869$ under the smart method. Since we have established that the longest queue is our front-runner for the optimization, we shall also use it as a basis for comparison for this smart method. Figure \ref{fig:longvsmart} contains the comparison between the average queue lengths for these two methods.

@@

\caption{Longest Queue (Left) vs. Smart Method (Right) average queue lengths}

@@

As expected based on the results of the lazy pharmacist method, we see improvement in the size of the verification queue lengths at the slight cost to the remainder of the queue lengths. Looking into the data we find,
\[\overline{Length}_{Longest}=1.389 > 1.187 = \overline{Length}_{Smart}\]
\[\overline{Throughput}_{Longest}=71.813 > 64.920 = \overline{Throughput}_{Smart}\]
It would appear that we have achieved a slight increase in effectiveness with this new implementation, but it is hard to say given the relative magnitudes of the standard deviations in the data, thus further analysis is required.

@@

Given that a pharmacy is a business above all else, the goal of these optimizations should be to increase the number of orders filled and reduce the employee idle times as much as possible. As such, we shall conduct a one-way ANOVA test on the mean number of orders filled to find the most effective model under the given conditions. Looking first at the IV filled data across the different methods, we see no marked difference among the four, they are all within one standard deviation of each other. As such, we shall focus on the oral filled data, as there is more significant variation in the results. First, let us consider a collation of the oral filled data across the methods, Table \ref{table:collate}.

@@

Lazy Pharmacist & $61.7$ & $4.85$ & $30$\\\hline

@@

Smart Method & $78.4$ & $6.04$ & $30$\\\hline

@@

We see from Figure \ref{fig:anovaBox} that there does seem to be a difference among the models, though it is hard to confirm visually, especially for the longest queue and smart methods. Running the ANOVA test on the null hypothesis of $\mu_i=\mu_j\ \forall i,j\in\text{Methods}$, we arrive at the $F$ statistic, 

@@

Lazy-Base   &  $-7.133333$ & $0.0001922$\\\hline

@@

Smart-Base  &   $9.600000$ & $0.0000003$\\\hline
Largest-Lazy & $11.633333$ & $0.0000000$\\\hline
Smart-Lazy  &  $16.733333$ & $0.0000000$\\\hline
\cellcolor{yellow!25}Smart-Largest & \cellcolor{yellow!25}$5.100000$ & \cellcolor{yellow!25}$0.0132541$\\\hline

@@

The results in Table \ref{table:tukey} are essentially pairwise $t$-tests to compare the means of each methods oral throughput. Note, Table \ref{table:tukey} has two rows highlighted. These rows correspond to the differences between the base method and largest queue method, as well as the largest queue method to the smart method. Using our standard $\alpha=0.05$, we can draw the following conclusion comparing the different methods,
\[\text{Smart $>$ Largest $>$ Base}\]
Note, this alpha comparison does not require a Bonferroni correction, as $R$ computes the adjusted $p$-value to correct for the multiple comparisons taking place. From these results, we see that while the largest queue did provide significant improvement over the base case, the smart method provided a significant improvement over the largest method. As such, we shall treat the smart method as the ``optimal" method for the next stage of our optimization efforts, improving staffing efficiency.

@@

While this may be the current conditions of the pharmacy, it is still possible that this is not the most cost effective staffing. In order to determine the best possible configuration of employees, we shall construct a $10\times 10$ mesh that defines the pharmacies profit as a function of the number of employees of each type. Here, each point in our mesh corresponds to a coordinate in the Technician-Pharmacist plane that has a magnitude of profit. In order to assure fairness, the simulation is run with the same time distribution sampling inputs for each coordinate configuration, and run 30 times to allow for random variance in choosing queues to be accounted for. Let us define the profit function as,

@@

Using the Smart Method developed earlier, we evaluate $\mathcal{P}$ over the mesh of staffing ``coordinates" in order to create a 3-D surface that describes the profit. This surface is then projected onto a 2-D heat-map weighted by profit in Figure \ref{fig:heatmap}.

@@

\caption{Heatmap of profit values using the Smart Method. Note, labeling on the $x$ and $y$ axes is reduced by one due to indexing.}

@@

As evidenced by the heat map, we see that the most profitable configurations occur for low numbers of pharmacists and higher numbers of technicians to compensate.  In fact, we find that the optimal number of employees for this test is one pharmacist and six technicians (5 normal and 1 IV). This resulted in an estimated net profit of $\mathcal{P}(1,6)=\$4894.02$ per day, though most of the surrounding mesh points had an approximately similar value, most specifically when moving along the technician axis.\\
This being said, there are a number of potential issues with this efficiency mesh. It makes broad assumptions about the value of each filled order, as well as approximate salary data. These approximations could be improved using real world data that the pharmacy would be able to provide, as well as the addition of specific salary data. This being said however, due to the massive discrepancy between the salary of pharmacists as opposed to technicians, it is unlikely that the ideal configuration will deviate all that much from our approximation.

@@

After performing both layers of optimization, we discovered a number of different effects that simple changes to task logic and staffing caused. While altering the task choice logic for the pharmacists and main technicians had dramatic effects on the overall effectiveness of the pharmacy, we noted that there was a minimal effect on the IV Technicican. In addition, we noticed that regardless of task prioritization method, as overall efficiency increased, the idle times of all employees decreased, while maintaining the trend that pharmacists had less idle time than technicians. This is likely due to the fact that they have a larger range of tasks to pull from, as well as the chance to steal a task from a technician.

@@

During the construction of our models, we used small sample sets of time data to estimate population distribution parameters. As such, we potentially ran the risk of under fitting the data, or estimating from fringe data sample cases. Larger samples of time data would help us to more accurately get a true estimation of the actual time population parameters.  Additionally, our model does not allow for employee skill and experience to be factored into the processes at all. It seems likely that a pharmacist would have more experience at performing technician level tasks, as they likely were employed as technicians before becoming pharmacists. This experience likely translates into faster task completion times, which our model does not account for. In this same vein, it is also likely that the incoming order rate is not a constant distribution, but rather a time sensitive process. Given that the Mayo Clinic has three main campuses in differing regions of the country, the levels of prescription needs are likely quite different. It is likely that more prescriptions are filled during the winter months in the Minnesota area due to an increase in hospitalizations due to vehicle accidents during inclement weather, as well as cold weather related sicknesses; whereas in Florida, there are probably more medications filled for long term conditions like hypertension and blood pressure management due to an older population base, and less winter effects to consider. As such, data specific to each region should be included into the model for optimal results.

@@

In addition to the potential issues with the time distribution estimates, there are also a number of human factors that were omitted for simplicity in this study that could potentially be addressed with more data. These factors include,

>>>>

The use of prescription medication is becoming increasingly commonplace in the United States. In 2013, the average number of prescriptions taken per person was $12.2$, up from $9$ in 1998, a $136\%$ increase over 15 years \cite{georgetown}\cite{statistica}. With this trend in prescription usage continuing to rise, pharmacies are becoming an ever greater presence in the lives of most Americans. As such, pharmacies need to adapt to deal with the increasing numbers of orders, as well as improving the speed and accuracy at which they fill prescriptions in order to remain competitive in an increasingly saturated landscape of pharmacies.  Consequently, pharmacies must adapt their methodologies and staffing requirements in order to optimize efficiency and cost. Testing different optimization schemas is both expensive and time consuming, so pharmacies have begun to consider computer simulation approaches to optimization in search of more cost effective solutions.\\\hfill\\
In this case study, we are tasked with both modeling and optimizing the workflow for a prototype pharmacy based on samples of data regarding task duration and incoming order frequency. In addition to the sample data for the pharmacy, we are also given a brief overview of the basic order fulfillment process. Using this information, we shall construct a discrete event simulation model to analyze our so called ``control" case, i.e. the pharmacy described by the basic information. Once this simulation is running and tested, we shall attempt to improve the efficiency over the base case. In order to achieve this goal, we shall attack the optimization on two fronts:

@@

With these optimizations, we will develop a sense of what changes could increase the efficiency of the pharmacy. We will also address which variables could be manipulated to further increase the accuracy of our simulation and therefore improve the pharmacy's function.

@@

In our prototypical example, there are two types of prescription orders that the pharmacy receives, oral and IV, which are handled by two different classes of employee, pharmacists and technicians, who are constrained as to which task they may perform due to legal regulations. In our prototype pharmacy, the proposed process is as follows: As the pharmacy receives orders, they are entered into the system by a technician. Once the order is in the system, it is verified by a pharmacist, and passed to the appropriate area to be prepared by technicians. Upon being filled, the order is again verified by a pharmacist, and then dispensed to the consumer by a technician. As such, we see that there are six basic tasks in the workflow tied to the two different types of orders: order entry, entry verification, oral or IV preparation, prepared order verification, and distribution through the window. As noted before, regulations exist to limit which tasks non-pharmacist level employees may perform. These regulations dictate that the verification tasks may only be performed by the pharmacists, while the remaining tasks can be performed by both pharmacists and technicians. A breakdown of this hierarchy follows in Table \ref{table:tasks}, with pharmacist specific tasks highlighted in red. It should also be noted that in our control case, there exists an additional sub-type of technician, whose sole purpose is to prepare IV orders in an isolated room. This will be explored in greater detail below.

@@

\begin{tabular}{|c||c||c|}

@@

Order Entry & Order Entry & \cellcolor{black}\\\hline
\cellcolor{red!25}Entry Verification &\cellcolor{black}& \cellcolor{black}\\\hline

@@

\cellcolor{red!25}Preparation Verification &\cellcolor{black}&\cellcolor{black}\\\hline
Distribution & Distribution&\cellcolor{black}\\\hline

@@

With this task breakdown in hand, we consider a graphical representation of the path an order takes within the pharmacy, specifically linked to the tasks described above. Looking at Figure \ref{fig:flowchart} below, we see that the process of filling an order is fairly linear, with the only fork in the workflow occurring after the entry verification step. This fork is due to the necessity of establishing an order's type (Oral vs. IV) so it may be put in the appropriate filling queue. In our control pharmacy, all IV orders are filled in an isolated room by an ``IV Technician" who only prepares IV prescriptions and does not assist with any other tasks in the workflow. 

@@

\caption{Pharmacy workflow by task position showing which tasks are pharmacist only.}

@@

For our control case, we operate under the assumption that no tasks in this process have any issues with failures or delays. This is a simplified case, but we shall construct the simulation in such a way that including such issues would be quite easy to do.  With this workflow defined, we consider the next step of building our simulation: identifying the durations of the different tasks.

@@

As noted above, there are six pivotal task durations that we must establish to effectively perform an analysis. In addition to these tasks, we are also aware of the fact that the times between incoming oral and IV orders are from different time distributions. As part of the supplied data, we have samples of the completion times of each of these eight different categories, ranging in size from 34 to 60 observations. Using each of these samples, we shall construct Cullen and Frey graphs to isolate the correct distributions. These graphs will take the skewness and kurtosis of our task time samples and plot a point in the $skew^2/kurtosis$ plane which can then be mapped to potential distributions based on theoretical values for each distribution. With this set of hypothetical distributions in hand, we shall fit the data and test the goodness of fit statistic to identify the best fit.

@@

Looking at the histogram of order entry data in Figure \ref{fig:orderEntryHistCullen}, it appears that the data comes from a normal population, and the Cullen and Frey plot supports this. For completeness, we shall test this sample as being from a normal distribution as well as a gamma and log-normal distribution, as these distributions also appear to be good potential fits in the Cullen and Frey graph. Once the data has been fit with each of these distributions, $R$ is used to compute a goodness of fit statistic, the AIC, for each model. The AIC is a numerical score for a model, which weights the accuracy of a fitted model against its own complexity. By construction, this score decreases for better models, and increases when models become too complex or over-fit. This scoring helps to reduce reliance on sample data values and many variable  models that do not capture the nature of the true population. With this in mind, we consider the AIC scores for each of the three potential distributions for order entry below in Table \ref{table:entryGOF}

@@

& Normal Fit & Gamma Fit & Log-Normal Fit \\\hline

@@

As highlighted in Table \ref{table:entryGOF}, by treating this sample as part of a larger normal population, we produce the smallest AIC score, implying the best fit. As such, we shall treat this as the correct distribution for sampling purposes in our simulation. The relevant distribution statistics follow:

@@

To construct the simulation of the pharmacy, the programming language C++ was used due to its speed and object oriented structure. This structure allows for maximal customization of our data, enabling the collection of many different statistics, which we report below. To build the simulation, two unique class objects were written to house the details required for a given prescription, as well as each of the employees. Within these classes relevant structures for computing the idle times and other important statistics are included and updated throughout the simulation. An outline of the pseudo-code for the simulation follows:

@@

It is important to note here the job assignment of the base case. In this model, when job assignment begins for each employee, each non-empty task queue is put into a pool and randomly chosen with no priority. So, a pharmacist has an equal probability of doing a pharmacist only task or a technician level task. Changing this assignment will become a cornerstone of our optimization efforts. To further remove any bias in the model, the idle worker pool is shuffled randomly before each task assignment step in the simulation. This will help to alleviate any artificial bias on worker idle times.

@@

We see that we expect just shy of $100$ oral prescriptions and $50$ IV orders per day. This indeed aligns with the expected values given the exponential distribution fit that was assigned to them for time between orders.\\\hfill\\
Next, we consider the average queue lengths for each task in the workflow:

@@

Next, we consider arguably the most valuable statistic to a pharmacy from a business standpoint: the number of filled orders. 

@@

We see here that we expect to fill about $70\%-75\%$ of all incoming orders. To see whether there is room for improvement in this number, we consider whether the amount of time required to fill orders from start to finish, the ``throughput time", as well as the amount of time the order sits idle can be optimized.

@@

While knowing how long and how fast the orders are moving in the model is important, it is equally important for the pharmacy to measure the efficiency and utilization of its employees. Below, we see the estimations for how long each type of employee is idle during the simulations, expressed as a percent of the total time.

@@

We see that the results of this simulation point to the verification queues as being congestion points. Notably, these are the two tasks in the workflow that are pharmacist specific. This congestion is likely due to the method by which workers received their job assignments. With the pharmacists choosing from all possible tasks equally, they can take work from the technicians, who cannot in return work from the verification task pools. This aligns with the estimated idle times from above, as we found the pharmacists to be idle significantly less than the technicians. Reducing this congestion will be the goal of our first optimization.

@@

With the base case of the model constructed and analyzed, we begin the first of our two pronged optimization efforts, optimizing the job choice logic to increase throughput speed and reduce idle time in the model.
\subsection*{First Optimization: Focused Pharmacists}
\addcontentsline{toc}{subsection}{Focused Pharmacists}
Given that the verification tasks are causing congestion in the base model, we consider the effects of changing the job assignment method in the simulation to the following:

@@

Under this scheme, the pharmacists will only do pharmacist specific tasks, leaving the technicians to do the remainder of the tasks. The hope is that removing the pharmacists from the technician tasks will reduce not only the technician idle times, but also the lengths of the verification queues. Running the simulation with the same parameters as the base case, but with the new job choice logic yields some unexpected results. Firstly, the average queue lengths are much different. Figure \ref{fig:basevprof} shows the differing results.

@@

\caption{Base case (left) vs. Focused Pharmacist (Right) average queue lengths}

@@

Under the original schema, all queues were essentially empty except for the verification queues, while under our new schema all of the non-verification queues show increases, most specifically the first and last steps of the process. Indeed we find that the average queue length for our new case is slightly larger than before,
\[\overline{Length}_{Base}=2.648<2.898=\overline{Length}_{Focused Pharmacist}\]

@@

\[\overline{Idle}_{Base}=14.475\% < 21.627\%=\overline{Idle}_{Focused Pharmacist}\]
Looking more closely into this idle time data, we see that the average idle time for a pharmacist in the new scheme is approximately $67\%$, which is still not an effective use of their time. Before looking to see if reducing the number of pharmacists in the model would help to reduce this idle percentage, we first consider the filled order results under this scheme. Figure \ref{fig:baseprofFill} shows us a comparison in throughput.

@@

\caption{Base case (left) vs. Focused Pharmacist (Right) average number of filled orders}

@@

We see that in fact this method has reduced both the number of oral and IV orders that are filled over the duration of the simulation, which is our most valuable statistic from a business standpoint. It is likely not worth exploring further in terms of staffing optimizations as this ``base-staffed" run yielded worse results than a method based purely on randomness. As such, we seek to find a better process optimization.

@@

Given that forcing the pharmacists to only pull from the verification tasks is hurting the efficiency of the model, we consider a new job assignment logic: choosing the most full task queue. In this optimization a pharmacist will choose to take their next task from the largest of all five possible queues, while technicians choose from the three technician level queues. This method should help to ensure that both employees and orders are moving as efficiently as possible because it favors maintaining the smallest possible net queue sizes. Running the simulation with the same parameters as the base case, under the new job choice logic yields results that are more aligned with what we would expect. Figure \ref{fig:basevlong} shows the comparison of average queue lengths between the base case and the longest queue method.

@@

We see that under this new system, all of the queues (except IV fill) have non-zero means, but that they are all relatively small compared to the mean of approximately $10$ orders for the base entry verification length. Looking deeper into the data, we see that the average queue length is reduced from $2.648$ orders in the base case to $1.389$ orders in this new method. In addition to this, average idle time for the workers is reduced from $14.475\%$ in our base model to $13.395\%$. While these changes may seem somewhat small, consider now the mean throughput times between the two models, as shown in Figure \ref{fig:basevlongthrough}.

@@

We see that by choosing the longest queue as the next task, the average throughput time for an order is reduced by a factor of almost $30\%$. As we would expect, this also helps to reduce the average wait time for orders, as well as the idle times of the employees.  While this method is clearly superior to the base case, we will consider finally a combination of the Focused Pharmacist and Longest Queue methods.
\subsection*{Third Optimization: Combined Method}
\addcontentsline{toc}{subsection}{Combined Method}
As our final job assignment optimization we consider a mixture between choosing the longest queue and relegating the pharmacists to the verification queues. Namely, under this schema, the pharmacist will first check to see if any of the verification queues are non-empty, and if so, they will choose the largest one. Otherwise, they choose the largest from the remaining queues. The technicians remain unchanged from before, always choosing a task from the largest technician queue.

@@

Again we implement this model with the basic parameters, getting better results than before. We consider the comparison between the base case and the combined method queue lengths to begin to test whether we have achieved optimization.

@@

\caption{Base Case (left) vs. Combined Method (Right) average queue lengths}

@@

From Figure \ref{fig:basevsmart}, we see a significant reduction in the average queue length, going from $2.6475$ orders in the base case to $1.1869$ orders  under the combined method. Since we have established that the longest queue is our front-runner for the optimization, we shall also use it as a basis for comparison for this combined method. Figure \ref{fig:longvsmart} contains the comparison between the average queue lengths for these two methods.

@@

\caption{Longest Queue (Left) vs. Combined Method (Right) average queue lengths}

@@

As expected based on the results of the focused pharmacist method, we see improvement in the size of the verification queue lengths at the slight cost to the remaining queue lengths. Looking into the data we find,
\[\overline{Length}_{Longest}=1.389 > 1.187 = \overline{Length}_{Combined}\]
\[\overline{Throughput}_{Longest}=71.813 > 64.920 = \overline{Throughput}_{Combined}\]
It would appear that we have achieved an increase in effectiveness with this new implementation, but it is hard to say given the relative magnitudes of the standard deviations in the data, thus further analysis is required.

@@

Given that a pharmacy is a business above all else, the goal of these optimizations should be to increase the number of orders filled and reduce the employee idle times as much as possible. As such, we shall conduct a one-way ANOVA test on the mean number of orders filled to find the most effective model under the given conditions. Looking first at the IV filled data across the different methods, we see no marked difference among the four, they are all within one standard deviation of each other. We shall focus on the oral filled data, as there is more significant variation in these results. First, let us consider a collation of the oral filled data across the methods, Table \ref{table:collate}.

@@

Focused Pharmacist & $61.7$ & $4.85$ & $30$\\\hline

@@

Combined Method & $78.4$ & $6.04$ & $30$\\\hline

@@

We see from Figure \ref{fig:anovaBox} that there does seem to be a difference among the models, though it is hard to confirm visually, especially for the longest queue and combined methods. Running the ANOVA test on the null hypothesis of $\mu_i=\mu_j\ \forall i,j\in\text{Methods}$, we arrive at the $F$ statistic, 

@@

Focused-Base   &  $-7.133333$ & $0.0001922$\\\hline

@@

Combined-Base  &   $9.600000$ & $0.0000003$\\\hline
Largest-Focused & $11.633333$ & $0.0000000$\\\hline
Combined-Focused  &  $16.733333$ & $0.0000000$\\\hline
\cellcolor{yellow!25}Combined-Largest & \cellcolor{yellow!25}$5.100000$ & \cellcolor{yellow!25}$0.0132541$\\\hline

@@

The results in Table \ref{table:tukey} are analogous to pairwise $t$-tests comparing the means of each method's oral throughput. Note: Table \ref{table:tukey} has two rows highlighted. These rows correspond to the differences between the base method and the largest queue method, as well as between the largest queue method and the combined method. Using our standard $\alpha=0.05$, we can draw the following conclusion comparing the different methods,
\[\text{Combined $>$ Largest $>$ Base}\]
Note: this alpha comparison does not require a Bonferroni correction, as $R$ computes the adjusted $p$-value to correct for the multiple comparisons taking place. From these results, we see that while the largest queue did provide significant improvement over the base case, the combined method provided significant improvement over the largest queue method. As such, we shall treat the combined method as the ``optimal" method for the next stage of our optimization efforts, improving staffing efficiency.

@@

While this may be the current conditions of the pharmacy, it is still possible that this is not the most cost effective staffing. In order to determine the best possible configuration of employees, we shall construct a $10\times 10$ mesh that defines the pharmacy's profit as a function of the number of employees of each type. Here, each point in our mesh corresponds to a coordinate in the Technician-Pharmacist plane that has a magnitude of profit. In order to assure fairness, the simulation is run with the same time distribution sampling inputs for each coordinate configuration, and run 30 times to allow for random variance in choosing queues to be accounted for. Let us define the profit function as,

@@

Using the Combined Method developed earlier, we evaluate $\mathcal{P}$ over the mesh of staffing ``coordinates" in order to create a 3-D surface that describes the profit. This surface is then projected onto a 2-D heat-map weighted by profit in Figure \ref{fig:heatmap}.

@@

\caption{Heatmap of profit values using the Combined Method. Note, labeling on the $x$ and $y$ axes is reduced by one due to indexing.}

@@

As evidenced by the heat map, we see that the most profitable configurations occur for lower numbers of pharmacists and higher numbers of technicians to compensate.  In fact, we find that the optimal number of employees for this test is one pharmacist and six technicians (5 normal and 1 IV). This resulted in an estimated net profit of $\mathcal{P}(1,6)=\$4894.02$ per day, though most of the surrounding mesh points had an approximately similar value, most specifically when moving along the technician axis.\\
This being said, there are a number of potential issues with this efficiency mesh. It makes broad assumptions about the value of each filled order, as well as approximate salary data. These approximations could be improved using real world data that the pharmacy would be able to provide, as well as the addition of specific salary data. However, due to the massive discrepancy between the salary of pharmacists as opposed to technicians, it is unlikely that the ideal configuration will deviate much from our approximation.

@@

After performing both layers of optimization, we discovered a number of different effects that simple changes to task logic and staffing caused. While altering the task choice logic for the pharmacists and main technicians had a dramatic impact on the overall effectiveness of the pharmacy, we noted that there was a minimal effect on the IV Technicican. In addition, we noticed that regardless of task prioritization method, as overall efficiency increased, the idle times of all employees decreased while still maintaining the trend that pharmacists had less idle time than technicians. This is likely due to the fact that they have a larger range of tasks to pull from as well as the chance to steal a task from a technician.

@@

During the construction of our models, we used small sample sets of time data to estimate population distribution parameters. As such, we potentially ran the risk of under-fitting the data or estimating from fringe data sample cases. Larger samples of time data would help us more accurately get an estimation of the actual time population parameters.  Additionally, our model does not allow for employee skill and experience to be factored into the processes. It seems likely that a pharmacist would have more experience at performing technician level tasks, as they likely were employed as technicians before becoming pharmacists. This experience likely translates into faster task completion times, which our model does not account for. In this same vein, it is also likely that the incoming order rate is not a constant distribution, but rather a time sensitive process. Given that the Mayo Clinic has three main campuses in differing regions of the country, the levels of prescription needs are likely quite different. It is likely that more prescriptions are filled during the winter months in the Minnesota area based on an increase in hospitalizations due to vehicle accidents during inclement weather as well as cold weather related sicknesses; whereas in Florida, there are probably more medications filled for long term conditions like hypertension and blood pressure management due to an older population base. As such, data specific to each region should be included into the model for optimal results.

@@

In addition to the potential issues with the time distribution estimates, there are also a number of human factors that were omitted for simplicity in this study that could potentially be addressed with more data. These factors include:
